Jacobi Elliptic Functions in $\RR\times\RR$

### 2016-03-22 pde_two/elliptic.tex
finished elliptic

\documentclass[notitlepage]{hw}
\title{Elliptic Functions}
\date{}
\author{}

\usepackage{pgf,tikz}
\usepackage{mathrsfs}
\usetikzlibrary{arrows}

@@

We will begin our discussion by considering an ellipse centered about the origin in the cartesian plane.
Such an ellipse can be given by the equation
\[
\left({x^{2}\over a^{2}}\right) + \left({y^{2}\over b^{2}}\right) = 1.
\]
It will benefit us to consider the case where $b=1$, which gives the equation
\[
\left({x^{2}\over a^{2}}\right) + \left(y^{2}\right) = 1.
\]
We can take a given radius $r$ at the point $Q$ on the ellipse, and a given point $P$ at $a$. In
addition, we can call the angle between the the $x$ axis and the radius $\theta$. Rather than consider
a point $(x,y)$, we will actually consider a point $Q=(x(u),y(u))$, where $u$ is a function that we will
define later. We can define the JEFs in terms of this ellipse; we have
\[
\begin{cases}
cn(u,k)=y\\
sn(u,k)={x\over a}\\
dn(u,k)={r\over a}
\end{cases}
\]
\begin{center}
\includegraphics[scale=0.5]{ellipsis}
\end{center}
We can establish some useful identities from this relationship, such as
\[
sn^{2}(u,k)+cn^{2}(u,k)=\left({x\over a}\right)^{2}+y^{2} = 1.
\]
This is similar to our trig identity of $\sin^{2}{x}+\cos^{2}{x}=1$. We also have that
\[
k^{2}sn^{2}(u,k)+dn^{2}(u,k)=1,
\]
which may not be as familiar, but it still holds true. Interestingly enough, we can formulate the
special (or maybe not so special) system of ODEs from the previous section using this new definition.

@@

There is a third definition of JEFs that we can consider; that is, the form they take from an integral.
Consider the ODE
\[
{\dx\over\dt}=\sqrt{(1-x^{2})(1-k^{2}x^{2})}.
\]
We can attempt separation of variables as an approach to solve this. Doing so yields
\[
\int_{0}^{sn(t,k)}{\dx\over\sqrt{(1-x^{2})(1-k^{2}x^{2})}}=t.
\]
The upper limit comes from the comparison to the integral definition of $\sin$:
\[
t=\int_{0}^{\sin{t}}{\dx\over\sqrt{1-x^{2}}}.
\]

\end{document}

>>>>

\documentclass[notitlepage]{hw}
\title{Jacobi Elliptic Functions in $\RR\times\RR$}
\date{}
\author{}

@@

We will begin our discussion by considering an ellipse centered about the origin in the cartesian plane.
Such an ellipse can be given by the equation
\[
\left({x^{2}\over a^{2}}\right) + \left({y^{2}\over b^{2}}\right) = 1.
\]
It will benefit us to consider the case where $b=1$, which gives the equation
\[
\left({x^{2}\over a^{2}}\right) + \left(y^{2}\right) = 1.
\]
We can take a given radius $r$ at the point $Q$ on the ellipse, and a given point $P$ at $a$. In
addition, we can call the angle between the the $x$ axis and the radius $\theta$. Rather than consider
a point $(x,y)$, we will actually consider a point $Q=(x(u),y(u))$, where $u$ is a function that we will
define later. We can define the JEFs in terms of this ellipse; we have

\vspace{0.5cm}

\begin{minipage}{0.5\textwidth}
\[
\begin{cases}
cn(u,k)=y\\
sn(u,k)={x\over a}\\
dn(u,k)={r\over a}
\end{cases}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\includegraphics[scale=0.5]{ellipsis}\\

A generalized ellipse
\end{center}
\end{minipage}

\vspace{0.5cm}

We can establish some useful identities from this relationship, such as
\[
sn^{2}(u,k)+cn^{2}(u,k)=\left({x\over a}\right)^{2}+y^{2} = 1.
\]
This is similar to our trig identity of $\sin^{2}{x}+\cos^{2}{x}=1$. We also have that
\[
k^{2}sn^{2}(u,k)+dn^{2}(u,k)=1,
\]
which may not be as familiar, but it still holds true. Interestingly enough, we can formulate the
special (or maybe not so special) system of ODEs from the previous section using this new definition.

\section{Properties of Jacobi Functions}

Before we continue, we will want to outline some fundamental properties of the JEFs. Recall that we
already saw the analog to $\sin^{2}{x} + \cos^{2}{x} = 1$, which is
\begin{gather}
sn^{2}(u,k)+cn^{2}(u,k)= 1.
\end{gather}
In addition, we have
\begin{gather}
k^{2}sn^{2}(u,k)+dn^{2}(u,k)=1.
\end{gather}
We can also related the derivatives of the JEFs by
\begin{align}
{d\over\dt}sn(t,k)&=cn(t,k)dn(t,k)\\
{d\over\dt}cn(t,k)&=-dn(t,k)sn(t,k)\\
{d\over\dt}dn(t,k)&=-k^{2}sn(t,k)cn(t,k).
\end{align}
Since the functions are differentiable, we can talk about the limits of JEFs. If we let $k\to0$, we
can see that our defining system of ODEs becomes
\[
\begin{cases}
\dot{x}=yz\\
\dot{y}=-zx\\
\dot{z}=0
\end{cases},
\]
which yields a basis of $\{\sin{t},\cos{t},1\}$, implying that
\begin{align}
\lim_{k\to0^{+}}sn(t,k)&=\sin{t}\\
\lim_{k\to0^{+}}cn(t,k)&=\cos{t}\\
\lim_{k\to0^{+}}dn(t,k)&=1.
\end{align}
We can also see what happens a $k\to1$:
\begin{align}
\lim_{k\to1^{-}}sn(t,k)&=\tanh{t}\\
\lim_{k\to1^{-}}cn(t,k)&=\sech{t}\\
\lim_{k\to1^{-}}dn(t,k)&=\sech{t}.
\end{align}
These are enough

@@

There is a third definition of JEFs that we can consider; the form they take from an integral. As we
saw previously, JEFs served as solutions to the system
\[
\begin{cases}
\dot{x}=yz\\
\dot{y}=-zx\\
\dot{z}=-k^{2}xy
\end{cases},
\]
but this is not the only system for which JEFs provide a solution. Consider the following system
\[
\begin{cases}
\dot{x}^{2}=(1-x^{2})(1-k^{2}x^{2})\\
\dot{y}^{2}=(1-y^{2})(\kappa^{2}+k^{2}y^{2})\\
\dot{z}^{2}=(1-z^{2})(z^{2}-\kappa^{2})
\end{cases}.
\]
We can transform the first ODE into
\[
{\dx\over\dt}=\sqrt{(1-x^{2})(1-k^{2}x^{2})},
\]
which is a separable equation. We can rewrite this as
\[
\int{\dx\over\sqrt{(1-x^{2})(1-k^{2}x^{2})}}=t,
\]
Which we may notice as an Elliptic Integral of the first kind. If we make the substitution
$x=\sin{u}$, then we have
\[
\int{\du\over\sqrt{1-k^{2}\sin^{2}{u}}}.
\]
This integral is not solvable by elementary means; rather we define the solution to be dependent on
$sn(u,k)$. There are similar integral definitions corresponding to $cn(u,k)$ and $dn(u,k)$.

\section{Applications to Elliptic Integrals}

We saw previously that we cannot integrate
\[
\int{\dt\over\sqrt{1-k^{2}\sin^{2}{t}}}
\]
using elementary methods. We saw that this integral came from a substitution on
\[
\int{\dx\over\sqrt{(1-x^{2})(1-k^{2}x^{2})}}.
\]
Recall that by property $(3)$, we have
\[
{d\over\dt}sn(t,k)=cn(t,k)dn(t,k)=\sqrt{(1-sn^{2}(t,k))(1-k^{2}sn^{2}(t,k))}
\]
by property $(1)$. This leads us to the result that
\[
\int{d\over\dt}sn(t,k)=sn(t,k)=\int\sqrt{(1-sn^{2}(t,k))(1-k^{2}sn^{2}(t,k))}\dt.
\]
Note that we can invert the $sn(t,k)$, which leads us to
\[
sn^{-1}(t,k)=\int{\dt\over\sqrt{(1-t^{2})(1-k^{2}t^{2})}}.
\]
We can similarly formulate definitions for $cn^{-1}(t,k)$ and $dn^{-1}(t,k)$.
\end{document}
